Local real space operators for translating Stokes Q/U to E/B maps on the sphere

### 2017-09-14 article/sections/discussion.tex
updated to conclusions and numerical implementation section

\item Small field experiments like BICEP, implicitly implement such radial cut offs due to limited survey area.

@@

>>>>

\item Small field experiments like BICEP implement such radial cut offs due to limited survey area.

@@

\item \textit{Total convolution methods:} Since the convolution kernels can be thought of as effective beams for polarization maps, it may be possible to use total convolution methods to construct E and B mode maps.

\item{Separate Q and U maps for E and B:} We have presented kernels which allow for decomposition of the total Stokes Q and U parameters to those corresponding to E and B modes. This could be potentially interesting, since one can now work with E and B modes foregrounds and their separations separately. \comment{But is there an issue with doing this in the standard method which involves going through the process of generating E and B modes ?}

### 2017-09-14 article/sections/numerical_implementation.tex
updated to conclusions and numerical implementation section

As discussed in \sec{sec:radial_locality}, the radial kernels decay ($\propto ~ \beta^{-1.5}$) as the distance from the central pixel increases, but they begin to increase on approaching the diametrically opposite position ($\beta \rightarrow \pi$)  as seen in \fig{fig:rad_ker_decay}). For a sufficiently high maximum multipole $\ell_{\rm max}$, the lowest multipole being fixed at $\ell_{\rm min}=2$, the maximum the radial kernel reaches near the vicity of the diametrically opposite end can be significantly smaller than global maxima of the radial kernel. This suggests that for sufficiently high resolution CMB maps (${\rm NSIDE} >32$), having high multipole information, it may suffice to restrict the convolution over the Stoke parameters Q \& U (to deduce either the scalar E \& B maps or pure Stokes Q \& U corresponding to either of the two scalar modes) to a local region around the central pixel. The locality of the region is quantified by the non-locality parameter $\beta_o$, which is a function of the maximum multipole accessible in the Stokes maps.

We have developed a Python script to evaluate these local convolution over Stokes Q \& U parameters. To evaluate the convolution kernel (\eq{eq:qu2eb_convolution}), we precompute the radial part of the kernel ( $f(\beta), {}_{+2}f(\beta) ~\&~ {}_{-2}f(\beta))$. Specifically, we use the python routine {\textit scipy.special.lpmn} to evaluate the associated Legendre polynomial functions, which are needed to evaluate the radial kernels. $f(\beta)$ is simply evaluated by calculating the multipole sum in \eq{eq:rad_ker_queb}. ${}_{+2}f(\beta) ~\&~ {}_{-2}f(\beta))$ are estimated by evaluating the sum in \eq{eq:f2_rad_ker}, where we take care to use the limiting forms of the functions ${}_{\pm 2}f_{\ell}(\beta)$ in the vicinity of $\beta=0,\pi$ as discussed in \app{sec:asymptotic_f}. As the next step, for each pixel on the Healpix map we get the pixel numbers of all the neighboring pixels lying within radius of $r_{\rm cutoff}$ from the central pixel using the Healpix routine {\textit query\_disc}. We then use the {\textit pix2ang} function of Healpix to get the angular coordinates of the central $(\theta_o,\phi_o)$ and its surrounding pixels $(\theta_i,\phi_i)$, which are then used to calculate the corresponding Euler angles using \eq{eq:fn_euler}. Given these  inputs we evaluate the convolution kernel as simple Reimann sums. We repeat this procedure for each pixel on the Healpix map.

For the results presented in the following sub-sections we work with CMB maps at Healpix resolution of ${\rm NSIDE}=64$, unless specified otherwise. \revisit{We use the CMB spectra for some fiducial cosmology. This work  only deals with lensing B-mode spectra.} To understand the effects of restricting the convolution to a local neighbourhood, we evaluate these convolutions on discs with progressively smaller radii surrounding the central pixel.  Specifically the non-locality parameter for a NSIDE=64 map is $\beta_o=30^{\circ}$. We impose radial cutoffs of $r_{\rm cutoff}=[2\beta_0,\beta_o,0.5\beta_o,0.25\beta_o]$ with an apodization of 3 degrees having a cosine squared profiles on the edges of the discs. We also evaluate the corresponding maps using standard Healpix routines and use these as reference maps for this exercise. Note that the Healpix evaluations are in principle equivalent to carrying out the convolution over the full sphere (i.e. $\beta_o=\pi$).  We compare the spectra derived from all these maps to those derived from the reference maps to quantify the effect of the imposed radial cutoff on the convolution integrals.

>>>>

As discussed in \sec{sec:radial_locality}, the radial kernels decay ($\propto ~ \beta^{-1.5}$) as the distance from the central pixel increases, but they begin to increase on approaching the diametrically opposite position ($\beta \rightarrow \pi$)  as seen in \fig{fig:rad_ker_decay}. For a map containing sufficiently high multipole $\ell_{\rm max}$ information,  the local maxima the radial kernel attains in the vicinity of the diametrically opposite end is significantly smaller than global maxima. This suggests that for sufficiently high resolution CMB maps, it may suffice to restrict the convolution over the Stoke parameters Q \& U to a local region around the central pixel. The size of this local region is determined by the non-locality parameter $\beta_o$. \revisit{It is important to note that this claim is valid only while working with maps that are fairly homogeneous. In the case of foregrounds for example, though the radial kernel falls off, a strong foreground very far from the central pixel may still make significant contribution to the local definition of E and B modes.}

In this section we study the effects of localizing the convolution kernels on the inferred E and B mode maps and their power spectra. We have developed a Python script to evaluate these local convolution over Stokes Q \& U parameters. To evaluate the convolution kernel (\eq{eq:qu2eb_convolution}), we precompute the radial part of the kernel ( $f(\beta), {}_{+2}f(\beta) ~\&~ {}_{-2}f(\beta))$. Specifically, we use the python routine {\textit scipy.special.lpmn} to evaluate the associated Legendre polynomial functions, which are needed to evaluate the radial kernels. $f(\beta)$ is simply evaluated by calculating the multipole sum in \eq{eq:rad_ker_queb}. ${}_{+2}f(\beta) ~\&~ {}_{-2}f(\beta))$ are estimated by evaluating the sum in \eq{eq:f2_rad_ker}, where we take care to use the limiting forms of the functions ${}_{\pm 2}f_{\ell}(\beta)$ in the vicinity of $\beta=0,\pi$, given in \app{sec:asymptotic_f}. As the next step, for each pixel on the Healpix map we get the pixel numbers of all the neighboring pixels lying within radius of $r_{\rm cutoff}$ from the central pixel using the Healpix routine {\textit query\_disc}. We then use the {\textit pix2ang} function of Healpix to get the angular coordinates of the central $(\theta_o,\phi_o)$ and its surrounding pixels $(\theta_i,\phi_i)$, which are then used to calculate the corresponding Euler angles using \eq{eq:fn_euler}. Given these  inputs we evaluate the convolutions as simple Reimann sums. We repeat this procedure for each pixel on the Healpix map.

\revisit{We use the CMB spectra for the fiducial cosmology. Specifically we restrict our analysis on lensing generated B-mode spectra.}  For the results presented in the following sub-sections we work with CMB maps at Healpix resolution of ${\rm NSIDE}=64$, unless specified otherwise. To understand the effects of restricting the convolution to a local neighbourhood, we evaluate these convolutions on discs with progressively smaller radii surrounding the central pixel.  Specifically the non-locality parameter for a NSIDE=64 map is $\beta_o=30^{\circ}$. We impose radial cutoffs of $r_{\rm cutoff}=[2\beta_0,\beta_o,0.5\beta_o,0.25\beta_o]$ with an apodization of 3 degrees having a cosine squared profiles on the edges of the discs. We also evaluate the corresponding maps using standard Healpix routines and use these as reference maps for this exercise. Note that the Healpix evaluations are equivalent to carrying out the convolution over the full sphere (i.e. $\beta_o=\pi$).  We compare the spectra derived from all these maps to those derived from the reference maps to quantify the effect of the imposed radial cutoff on the convolution integrals.
